---
title: "PA_template"
author: "Faisal Bin Basha"
date: "May 11, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(plyr)
library(lubridate)
library(ggplot2)
library(scales)
library(Hmisc)

```

# Loading and preprocessing the data
```{r}
activity <- read.csv("activity.csv",stringsAsFactors = FALSE,header=TRUE, sep=',',colClasses=c("numeric","character","integer"))
```

## R Markdown Reading the data into activity data frame and show some summary statistics

```{r}
activity <- read.csv("activity.csv",stringsAsFactors = FALSE)
summary(activity)
```

## Convert date to POSIXct clas using lubridate package
```{r}
activity$day <- weekdays(as.Date(activity$date))
activity$DateTime <- as.POSIXct(activity$date,format="%Y-%m-%d")

no_nas <- activity[!is.na(activity$steps),]
str(no_nas)

```
## head of no_nas
```{r}
head(no_nas)
```

# What is mean total number of steps per day

## calculate the total number of steps taken per day

```{r}
sumoftables <- aggregate(activity$steps ~ activity$date,FUN=sum,)

colnames(sumoftables) <- c("Date","Steps")

hist(sumoftables$Steps,breaks=5,xlab="Steps", main="Total Steps per Day")
```

## Calculate and report the mean and median of the total ## number of steps taken per day

```{r}
meanR <- as.integer(mean(sumoftables$Steps))
medianR <- as.integer(median(sumoftables$Steps))
```

### Mean : 
```{r} 
meanR
```

### Median : 
```{r} 
medianR 
```


## What is the average daily activity pattern?

```{r} 
library(dplyr)

interval <- activity %>% 
            filter(!is.na(steps)) %>%
            group_by(interval) %>%
            summarize(steps = mean(steps))

average <- ggplot(interval,aes(x=interval,y=steps))

average + geom_line(color = "blue") + xlab("Interval") + ylab("Average Number of Steps") + ggtitle("Average number of steps per Interval")
```

## Maximum steps per interval
```{r}
maxSteps <- max(interval$Avg)
```

## Which interval contains the maximum average number of ## steps

```{r}

interval[interval$Avg == maxSteps,1]
```
#Imputing missing values

## Calculate and report the total number of missing values  in the dataset

```{r}
nrow(activity[is.na(activity$steps),])
```

## Devise a strategy for filling in all of the missing 
## values in the dataset

```{r}
dataImputed <- activity
dataImputed$steps <- impute(activity$steps,fun=mean)
```

## Histogram of the total number of steps per day
```{r}
stepsPerDayImputed <- tapply(dataImputed$steps,dataImputed$date,sum)
qplot(stepsPerDayImputed,xlab="Total steps per day (Imputed)",ylab="Frequency using binwidth 500",binwidth=500)
```

## Calculate and report the mean and median total number of steps taken per day

```{r}
stepsPerDayMeanImputed <- mean(stepsPerDayImputed)
stepsPerDayMedianImputed <- median(stepsPerDayImputed)
```

# Mean (imputed): 
```{r} 
stepsPerDayMeanImputed 
```

# Median (imputed): 
```{r} 
stepsPerDayMedianImputed 
```


# Are there differences in activity patterns between 
# weekdays and weekends

### 1. Create a new factor variable in the dataset with 
### two levels - "weekday" and "weekend" indicating 
### whether a given date is a weekday or weekend day

```{r}
weekday.or.weekend <- function(date) {
    day <- weekdays(date)
    if (day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")) 
        return("weekday") else if (day %in% c("Saturday", "Sunday")) 
        return("weekend") else stop("invalid date")
}
dataImputed$date <- as.Date(dataImputed$date)
dataImputed$day <- sapply(dataImputed$date, FUN = weekday.or.weekend)
```

### 2 Make a panel plot containing a time series plot

```{r}
averages <- aggregate(steps ~ interval + day, data = dataImputed, mean)
ggplot(averages, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) + 
    xlab("5-minute interval") + ylab("Number of steps")
```

